A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data

### 2024-04-23 review/response.qmd

---
title: "Response to Reviewers JCGS-23-428"
subtitle: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
author: "H. Sherry Zhang, Dianne Cook, Ursula Laa, Nicolas Langrené, Patricia Menéndez"
date: "2024-03-31"
format: pdf
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
We would like to thank the reviewers and editor for taking the time to review our paper and provide constructive comments. Below is a point-by-point explanation of how we have changed the paper. Reviewers' comments are in black and **\textcolor{purple}{our responses are in red}**.

@@

# Editor and Associate Editor

Two referees, an Associate Editor (AE), and I have reviewed your paper. The two referees hold significantly different views on the paper. As you will see, the first referee's assessment is rather negative, whereas the second is more positive. As recommended by the AE, it is essential to incorporate or highlight your methodological contributions in the paper to address the concerns raised by the first referee. Therefore, I would like to extend an opportunity for revision. However, the decision for the revision will depend on whether you can adequately address the reviewers' concerns.

**\textcolor{purple}{We have made changes to the introduction and to the conclusions being more specific about the methodological and computational contributions of this work.}**

@@

# Reviewer: 2

This paper proposes new pipeline work for index development, including index generation, validation, and sensitivity analysis. The authors develop a new R package, "tidyindex" including all functions in the pipeline to produce a new index. This package will be beneficial for developing new indices. The central part of this paper is the development of the R package and the explanation of its usage. However, this package needs to be fixed.

The authors explain their functions briefly in Chapter 5 and use examples with R code in Chapter 6. However, the code in Chapter 6 does not work. Also, the demo in the R package (CRAN version 0.1.0) did not work (see attached R-output.html). Please check the package, including demos and examples in this paper. It would be great if all R codes for the example in Chapter 6 were provided, including generating plots.

**\textcolor{purple}{ This issue is now fixed in the Demo folder (the vignette folder worked fine) and we have also re-organised the code for the paper as follows:}**

- **\textcolor{purple}{The main file tidyindex.R, can be used to reproduce the paper from saved data objects, available in the data folder.}** 
- **\textcolor{purple}{New scripts can be found in the 'scripts' folder, which generate each of the examples. The README.md file has been updated and it contains instructions on how to set up the R environment to reproduce all of the analyses in the paper.}**
- **\textcolor{purple}{The GitHub repo is listed in the Acknowledgements section, and any further updates needed for future versions to run will be made there.}**

@@

## minor comments

- In Table 1, it would be great if there were a line between items. Because the name of the module is in two lines, it is not easy to match the other information (input, operation, and output)

**\textcolor{purple}{The table layout has been modified accordingly.}**

@@

 - On page 19, " The output contains the original data, index values (.index), parameters used (.scale,
.method, and .dist), and all the intermediate variables (.pet, .agg, and .fitted)." However, the output has no .index, .pet, .agg, and .fitted. Please check.

**\textcolor{purple}{ The output is now written in a way that all variables are visible (availble from dplyr::glimpse()).}**

@@

- Figure 7 shows the sequence of screenshots from the animation to explore the sensitivity of the GGGI. It would be helpful to understand "what Frame 12 is"  if the authors should explain this between " Figure 7 illustrates doing sensitivity analysis for GGGI, for a subset of 16 countries" and "Frame 12 shows the dot plot of the original index values sorted from highest to lowest",

**\textcolor{purple}{Frame 12 is now described in more detail, with additional explanations provided regarding the information conveyed by the figure.}**

@@

- Figure 8 on page 27 also shows the relatively wide confidence interval around 2004. It would be great if the author mentioned it, too.

**\textcolor{purple}{The following has been added in the text: This wider confidence interval is interesting and corresponds to the millennium drought in Australia.}**

### 2024-04-24 review/response.qmd

**\textcolor{purple}{We have made changes to the introduction and to the conclusions being more specific about the methodological and computational contributions of this work.}**

# Reviewer: 1

I commend the authors for their work in the tidy index library.  Certainly, a user friendly and unified approach to indices is useful across many fields.  My main concern about the paper is that it doesn't seem to fit within the aims and scope of JCGS.  A perusal of recent JCGS papers shows that a novel statistical component is needed for publication in addition to the software.  This paper, while certainly useful, doesn't necessarily propose new statistical methods but rather proposes a great tidy framework for handling index calculation.  As such, for JCGS, I think the authors need to highlight the statistical contributions more than the library.  However, I don't think the authors statistical contributions (even if highlighted) are sufficiently novel to warrant publication in JCGS.  As written, I think the paper is more suited for publication in, say, Journal of Statistical Software than JCGS.

**\textcolor{purple}{It is a bit disappointed to see the following comment ``A perusal of recent JCGS papers shows that a novel statistical component is needed for publication in addition to the software." We disagree with that statement based on the journal scope:  "The Journal of Computational and Graphical Statistics (JCGS) presents the very latest techniques on improving and extending the use of computational and graphical methods in statistics and data analysis." The contributions that this paper makes are in computational statistics for an especially prevalent data analysis activity, that also provides potential for making better graphics in this area. We have added language to the introduction and to the conclusions that hopefully helps see these contributions.}**

### 2024-04-24 review/response.qmd

**\textcolor{purple}{We have made changes to the introduction and to the conclusions being more specific about the methodological and computational contributions of this work.}**

# Reviewer: 1

I commend the authors for their work in the tidy index library.  Certainly, a user friendly and unified approach to indices is useful across many fields.  My main concern about the paper is that it doesn't seem to fit within the aims and scope of JCGS.  A perusal of recent JCGS papers shows that a novel statistical component is needed for publication in addition to the software.  This paper, while certainly useful, doesn't necessarily propose new statistical methods but rather proposes a great tidy framework for handling index calculation.  As such, for JCGS, I think the authors need to highlight the statistical contributions more than the library.  However, I don't think the authors statistical contributions (even if highlighted) are sufficiently novel to warrant publication in JCGS.  As written, I think the paper is more suited for publication in, say, Journal of Statistical Software than JCGS.

**\textcolor{purple}{It is a bit disappointing to see the following comment ``A perusal of recent JCGS papers shows that a novel statistical component is needed for publication in addition to the software." We disagree with that statement based on the journal scope:  "The Journal of Computational and Graphical Statistics (JCGS) presents the very latest techniques on improving and extending the use of computational and graphical methods in statistics and data analysis." The contributions that this paper makes are in computational statistics for an especially prevalent data analysis activity, that also provides potential for making better graphics in this area. We have added language to the introduction and to the conclusions that hopefully helps see these contributions.}**

### 2024-04-24 tidyindex.qmd
add missing references

---
title: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
format:
  tandf-pdf:
    keep-tex: true  
    linestretch: 2
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: University of Texas at Austin
        department: Department of Statistics and Data Science
        address: Welch 5.216, 105E 24th Street, Stop G3800
        city: Austin, Texas
        country: USA
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@austin.utexas.edu
  - name: Dianne Cook
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-3813-7155
  - name: Ursula Laa
    affiliation: 
      - name: University of Natural Resources and Life Sciences
        department: Institute of Statistics
        address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
        city: Vienna
        country: Austria
    orcid: 0000-0002-0249-6439
  - name: Nicolas Langrené
    affiliation: 
      - name: BNU-HKBU United International College
        department: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science
        address: 2000 Jintong Road, Tangjiawan, Zhuhai
        city: Zhuhai, Guangdong
        country: China
    orcid: 0000-0001-7601-4618
  - name: Patricia Menéndez
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
      - name: University of Melbourne
        department: School of Mathematics and Statistics
        address: 813 Swanston Street, Parkville VIC 3052, Australia
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0003-0701-6315
abstract: |
  Indexes are useful for summarizing multivariate information into single metrics for monitoring, communicating, and decision-making. While most work has focused on defining new indexes for specific purposes, more attention needs to be directed towards making it possible to understand index behavior in different data conditions, and to determine how their structure affects their values and the variability therein. Here we discuss a modular data pipeline recommendation to assemble indexes. It is universally applicable to index computation and allows investigation of index behavior as part of the development procedure. One can compute indexes with different parameter choices, adjust steps in the index definition by adding, removing, and swapping them to experiment with various index designs, calculate uncertainty measures, and assess indexes' robustness. The paper presents three examples to illustrate the usage of the pipeline framework: comparison of two different indexes designed to monitor the spatio-temporal distribution of drought in Queensland, Australia; the effect of dimension reduction choices on the Global Gender Gap Index (GGGI) on countries' ranking; and how to calculate bootstrap confidence intervals for the Standardized Precipitation Index (SPI). The methods are supported by a new R package, called `tidyindex`.
keywords:
  - indexes
  - data pipeline
  - software design
  - uncertainty
  - decision-making
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

@@

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue raised in @donoho_50_2017, that different data analysts might arrive at different conclusions despite using the same data. This is especially pertinent to index use which affects important decisions such as in natural disaster prevention, economic interventions, resource allocation or human development. It is primarily due to a lack of standards in data analysis workflow. @donoho_50_2017 called for research on structuring unified workflow to address methodological variation across studies in data science. Current practices also violate statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric, for example by incorporating bootstrap confidence intervals [@efron_bootstrap_1979]. There has been considerable research in tidying up routine data analyses [XXX tidyr, dplyr, tidymodels, ..., tsibble, cubble] that turn ideas into software quickly and faithfully, as envisioned by @chambers. Index development and use needs tidying. <!--In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. -->.

@@

# Reference

>>>>

---
title: "A Tidy Framework and Infrastructure to Systematically Assemble Spatio-temporal Indexes from Multivariate Data"
format:
  tandf-pdf:
    keep-tex: true  
    linestretch: 2
    fontsize: 12pt
pdf-engine: pdflatex
author:
  - name: H. Sherry Zhang
    affiliations:
      - name: University of Texas at Austin
        department: Department of Statistics and Data Science
        address: Welch 5.216, 105E 24th Street, Stop G3800
        city: Austin, Texas
        country: USA
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-7122-1463
    email: huize.zhang@austin.utexas.edu
  - name: Dianne Cook
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0002-3813-7155
  - name: Ursula Laa
    affiliation: 
      - name: University of Natural Resources and Life Sciences
        department: Institute of Statistics
        address: Gregor-Mendel-Straße 33, 1180 Wien, Austria
        city: Vienna
        country: Austria
    orcid: 0000-0002-0249-6439
  - name: Nicolas Langrené
    affiliation: 
      - name: BNU-HKBU United International College
        department: Guangdong Provincial Key Laboratory of Interdisciplinary Research and Application for Data Science
        address: 2000 Jintong Road, Tangjiawan, Zhuhai
        city: Zhuhai, Guangdong
        country: China
    orcid: 0000-0001-7601-4618
  - name: Patricia Menéndez
    affiliation: 
      - name: Monash University
        department: Department of Econometrics and Business Statistics
        address: 29 Ancora Imparo Way, Clayton
        city: Melbourne, Victoria
        country: Australia
      - name: University of Melbourne
        department: School of Mathematics and Statistics
        address: 813 Swanston Street, Parkville VIC 3052, Australia
        city: Melbourne, Victoria
        country: Australia
    orcid: 0000-0003-0701-6315
abstract: |
  Indexes are useful for summarizing multivariate information into single metrics for monitoring, communicating, and decision-making. While most work has focused on defining new indexes for specific purposes, more attention needs to be directed towards making it possible to understand index behavior in different data conditions, and to determine how their structure affects their values and the variability therein. Here we discuss a modular data pipeline recommendation to assemble indexes. It is universally applicable to index computation and allows investigation of index behavior as part of the development procedure. One can compute indexes with different parameter choices, adjust steps in the index definition by adding, removing, and swapping them to experiment with various index designs, calculate uncertainty measures, and assess indexes' robustness. The paper presents three examples to illustrate the usage of the pipeline framework: comparison of two different indexes designed to monitor the spatio-temporal distribution of drought in Queensland, Australia; the effect of dimension reduction choices on the Global Gender Gap Index (GGGI) on countries' ranking; and how to calculate bootstrap confidence intervals for the Standardized Precipitation Index (SPI). The methods are supported by a new R package, called `tidyindex`.
keywords:
  - indexes
  - data pipeline
  - software design
  - uncertainty
  - decision-making
bibliography: bibliography.bib  
editor_options: 
  chunk_output_type: console
---

@@

<!-- what is the gap - why is this statistics -->
The problem is that every index is developed in its own unique way, by different researchers or organizations, and often indexes designed for the same purpose cannot easily be compared. This echoes an issue raised in @donoho_50_2017, that different data analysts might arrive at different conclusions despite using the same data. This is especially pertinent to index use which affects important decisions such as in natural disaster prevention, economic interventions, resource allocation or human development. It is primarily due to a lack of standards in data analysis workflow. @donoho_50_2017 called for research on structuring unified workflow to address methodological variation across studies in data science. Current practices also violate statistical principles, where quantifying and understanding uncertainty are essential to deciding on a best measure or metric, for example by incorporating bootstrap confidence intervals [@efron_bootstrap_1979]. There has been considerable research in tidying up routine data analyses [@wickham_tidy_2014; @wickham_split-apply-combine_2011; @tidymodels; @wang2020; @zhang2024cubble] that "turn ideas into software quickly and faithfully", as envisioned by @chambers1998programming. Index development and use needs tidying. <!--In Greater Data Science (GDS) 6: Science About Data Science, Donoho envisions future research on structuring unified workflow to address methodological variation across studies in data science. -->

@@

# References

### 2024-04-24 tidyindex.tex
quote
add missing references

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
]{interact}

@@

The problem is that every index is developed in its own unique way, by
different researchers or organizations, and often indexes designed for
the same purpose cannot easily be compared. This echoes an issue raised
in Donoho (2017), that different data analysts might arrive at different
conclusions despite using the same data. This is especially pertinent to
index use which affects important decisions such as in natural disaster
prevention, economic interventions, resource allocation or human
development. It is primarily due to a lack of standards in data analysis
workflow. Donoho (2017) called for research on structuring unified
workflow to address methodological variation across studies in data
science. Current practices also violate statistical principles, where
quantifying and understanding uncertainty are essential to deciding on a
best measure or metric, for example by incorporating bootstrap
confidence intervals (Efron 1979). There has been considerable research
in tidying up routine data analyses {[}XXX tidyr, dplyr, tidymodels,
\ldots, tsibble, cubble{]} that turn ideas into software quickly and
faithfully, as envisioned by (\textbf{chambers?}). Index development and
use needs tidying. .

@@

Since the success of tidyverse, more packages have been developed to
analyze data using the tidy framework for domains specific applications,
a noticeable example of which is \texttt{tidymodels} for building
machine learning models (Kuhn and Wickham 2020). To create a tidy
workflow tailored to a specific domain, developers first need to
identify the fundamental building blocks to create a workflow. These
components are then implemented as modules, which can be combined to
form the pipeline. For example, in supervised machine learning models,
steps such as data splitting, model training, and model evaluation are
commonly used in most workflow. In the \texttt{tidymodels}, these steps
are correspondingly implemented as package \texttt{rsample},
\texttt{parsnip}, and \texttt{yardstick}, agnostic to the specific model
chosen. The uniform interface in tidymodels frees analysts from
recalling model-specific syntax for performing the same operation across
different models, increasing the efficiency to work with different
models simultaneously.

@@

\section*{Reference}\label{reference}
\addcontentsline{toc}{section}{Reference}

@@

@@

\bibitem[\citeproctext]{ref-tidymodels}
Kuhn, Max, and Hadley Wickham. 2020. \emph{Tidymodels: A Collection of
Packages for Modeling and Machine Learning Using Tidyverse Principles.}
\url{https://www.tidymodels.org}.

@@

\bibitem[\citeproctext]{ref-wickham_tidy_2014}
Wickham, Hadley. 2014. {``Tidy {Data}.''} \emph{Journal of Statistical
Software} 59 (September): 1--23.
\url{https://doi.org/10.18637/jss.v059.i10}.

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
]{interact}

@@

The problem is that every index is developed in its own unique way, by
different researchers or organizations, and often indexes designed for
the same purpose cannot easily be compared. This echoes an issue raised
in Donoho (2017), that different data analysts might arrive at different
conclusions despite using the same data. This is especially pertinent to
index use which affects important decisions such as in natural disaster
prevention, economic interventions, resource allocation or human
development. It is primarily due to a lack of standards in data analysis
workflow. Donoho (2017) called for research on structuring unified
workflow to address methodological variation across studies in data
science. Current practices also violate statistical principles, where
quantifying and understanding uncertainty are essential to deciding on a
best measure or metric, for example by incorporating bootstrap
confidence intervals (Efron 1979). There has been considerable research
in tidying up routine data analyses (Wickham 2014, 2011; Kuhn and Silge
2022; Wang, Cook, and Hyndman 2020; Zhang et al. 2024) that ``turn ideas
into software quickly and faithfully'', as envisioned by Chambers
(1998). Index development and use needs tidying.

@@

Since the success of tidyverse, more packages have been developed to
analyze data using the tidy framework for domains specific applications,
a noticeable example of which is \texttt{tidymodels} for building
machine learning models (Kuhn and Silge 2022). To create a tidy workflow
tailored to a specific domain, developers first need to identify the
fundamental building blocks to create a workflow. These components are
then implemented as modules, which can be combined to form the pipeline.
For example, in supervised machine learning models, steps such as data
splitting, model training, and model evaluation are commonly used in
most workflow. In the \texttt{tidymodels}, these steps are
correspondingly implemented as package \texttt{rsample},
\texttt{parsnip}, and \texttt{yardstick}, agnostic to the specific model
chosen. The uniform interface in tidymodels frees analysts from
recalling model-specific syntax for performing the same operation across
different models, increasing the efficiency to work with different
models simultaneously.

@@

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

@@

\bibitem[\citeproctext]{ref-chambers1998programming}
Chambers, John M. 1998. \emph{Programming with Data: A Guide to the s
Language}. Berlin, Heidelberg: Springer-Verlag.

@@

\bibitem[\citeproctext]{ref-tidymodels}
Kuhn, Max, and Julia Silge. 2022. \emph{Tidy Modeling with r}. "
O'Reilly Media, Inc.".

@@

\bibitem[\citeproctext]{ref-wang2020}
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data
Structure to Support Exploration and Modeling of Temporal Data.''}
\emph{Journal of Computational and Graphical Statistics} 29 (3):
466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\bibitem[\citeproctext]{ref-wickham_split-apply-combine_2011}
Wickham, Hadley. 2011. {``The {Split}-{Apply}-{Combine} {Strategy} for
{Data} {Analysis}.''} \emph{Journal of Statistical Software} 40 (April):
1--29. \url{https://doi.org/10.18637/jss.v040.i01}.

\bibitem[\citeproctext]{ref-wickham_tidy_2014}
---------. 2014. {``Tidy {Data}.''} \emph{Journal of Statistical
Software} 59 (September): 1--23.
\url{https://doi.org/10.18637/jss.v059.i10}.

@@

\bibitem[\citeproctext]{ref-zhang2024cubble}
Zhang, H. Sherry, Dianne Cook, Ursula Laa, Nicolas Langrené, and
Patricia Menéndez. 2024. {``Cubble: An r Package for Organizing and
Wrangling Multivariate Spatio-Temporal Data.''}
\url{https://arxiv.org/abs/2205.00259}.
